Biomass predictions and VOCC

### 2019-08-14 analysis/VOCC/4-vocc-by-spp.Rmd
automate optimum value selection and save coefs

Species
```{r global_options, include=FALSE}
library(dplyr)
library(ggplot2)
# library(gfplot)
# library(gfdata)
library(sdmTMB)
library(gfranges)
# 
# knitr::opts_chunk$set(
#   fig.width = 11, fig.height = 8.5,
#   echo = FALSE, warning = FALSE, message = FALSE
# )
```

```{r}
# # Species run so far...
 species <- "Arrowtooth Flounder"
# species <- "Pacific Cod"
# species <- "Sablefish"
# species <- "Silvergray Rockfish"
#  species <- "Lingcod"
#  species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
#  species <- "Quillback Rockfish"
#  species <- "Pacific Ocean Perch"
# species <- "Yelloweye Rockfish"
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
```

Region 
```{r}
# region <- "All synoptic surveys"
 region <- "West Coast Vancouver Island"
# region <- "Both odd year surveys"
# region <- "West Coast Haida Gwaii"
```

@@

#climate <- "temperature"
temp_threshold <- c(0.75)
optimal_temp <- 6.5 # yellow-eye 4

#climate <- "do"
do_threshold <- c(0.25)
optimal_do <- 2 # yellow-eye 4

climate <- "do and temperature"
max_thresholds <- c(Inf, 0.5)
min_thresholds <- c(0.1, Inf)
optimal_values <- c(2, 6.5)

# critical_thresholds <- c(-0.25, 0.75)

@@

Run all subsequent code...

```{r, message=FALSE, warning=FALSE}
if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  trend_indices_do <- c(1, 1, 1, 2, 2)
  trend_indices_temp <- c(1, 1, 1, 2, 2)
  years <- NULL
}

@@

```{r}
rm(vocc)
# try({
#   vocc <- readRDS(paste0("data/vocc-", ssid_string,"-", climate_string, "-", optimal_string, "-",
#     min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
# })

@@

saveRDS(vocc, file = paste0("data/vocc-", ssid_string,"-", climate_string, "-", optimal_string, "-",
    min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
}
# glimpse(vocc)
```

### Plot climate vectors
```{r}

@@

@@

# Extract biomass values to match spatial and temporal range of climate vectors
```{r}

# #   #climate <- "do and temperature"
# max_thresholds <- c(Inf, 0.5)
# min_thresholds <- c(0.1, Inf)
# optimal_values <- c(2, 6.5)

vectors <- trim_vector_data(vocc, variable_names, optimal_values, min_thresholds, max_thresholds)

if (length(vectors[ ,2]) == 0) {
  other_coefs <- readRDS(paste0("data/",climate_string,"-vocc-by-spp-sdmTMB-coefs.RDS"))
  coefs <- other_coefs[1,]
  coefs$var <- "NA"
  coefs$beta <- "NA"
  coefs$se <- "NA"
  coefs$model <- "NA"
  coefs$species <- spp 
  coefs$lifestage <- "NA"
  coefs$climate <- climate 
  coefs$min_thresholds <- min_string
  coefs$max_thresholds <- max_string
  coefs$optimum <- optimal_string
  coefs$start_time <- start_time
  coefs$timespan <- delta_t_total
  coefs$region <- region
  coefs$ssid_string <- ssid_string

  coefs <- rbind(other_coefs, coefs) %>% distinct()
  saveRDS(coefs, file = paste0("data/",climate_string,"-vocc-by-spp-sdmTMB-coefs.RDS"))

@@

biomass_change <- make_trend_data(i,
  ssid = model_ssid,
  start_time = start_time,
  skip_time = skip_time,
  end_time = end_time,
  input_cell_size = input_cell_size,
  scale_fac = scale_fac,
  delta_t_total = delta_t_total,
  delta_t_step = delta_t_step,
  indices = indices,
  variable_names = "est_exp"
)
# browser()
vectors <- vectors %>%
      mutate(
        start_cell = icell,
        vector_id = paste0(id, "_to_", tid),
        distance2 = distance*distance,
        source_end_value = e, # may need to turn into numeric vector later...
        target_value = target_values
      )

@@

# combine source and target cell data
vocc_by_sp <- rbind(vector_source, vector_target) %>% 
  mutate(cell_type = as.factor(cell_type)) %>%
  filter(!is.na(icell)) %>%

  # remove potential source cells with lowest 1% quantile of fish biomass  
  filter(source_density > quantile(source_density, 0.01) ) %>%

  # calculate percent change in biomass
  mutate(
    log_change = log(end_density/start_density),
    perc_change = end_density/start_density
    )

end_climate <- purrr::map_df(strsplit(vocc_by_sp$end_climate, " "), 
                   function(x) data.frame(t(as.numeric(x))))

@@

vocc_by_sp
})

# glimpse(vocc_by_sp)
# trim spatial outliers 
# Xmax <- quantile(vocc_by_sp[[i]]$X, 0.9995)
# Xmin <- quantile(vocc_by_sp[[i]]$X, 0.005)
# std_max <- quantile(vocc_by_sp[[i]]$std_change, 0.9995)
# 
# vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% filter ( X >= Xmin & X <= Xmax ) 
```

@@

Check final climate-biomass relationship
```{r}
data <- vocc_by_sp[["adult"]]
#glimpse(data)

#for (j in seq_along(variable_names)) {

mature_plot <- lapply(variable_names, function(j) { 
ggplot(data, aes(x = UQ(rlang::sym(paste0("end_", j))), y = end_density, colour = cell_type, size = start_density, 
  group = cell_type)) +
  geom_point(alpha = 0.5) +
  # geom_smooth(method="lm", aes(x = UQ(rlang::sym(paste0("end_", j))), y = log_change
  # ), formula = y ~ x + I(x^2), inherit.aes = FALSE ) + # + I(x^3)
  # scale_size_continuous(guide = "none") +
  scale_y_continuous(trans = fourth_root_power) +
  xlab(paste("Final value of", j,"for cells")) +
  ylab(paste("Final biomass")) +
  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
  ggtitle(paste(survey, "mature", species))
})
print(mature_plot)
```

Check distance threshold of 50 looks reasonable
```{r}
data <- vocc_by_sp[["adult"]]
#data <- filter(vocc_by_sp[["adult"]],  distance < 90) # distance > 1 &
#data <- filter(vocc_by_sp[["adult"]], distance > 1 ) #  &

mature_plot2 <- lapply(variable_names, function(j) { 
  ggplot(data, aes(x = distance, y = log_change, 
      group = cell_type, size = start_density, 
      colour = UQ(rlang::sym(paste0("end_", j))))) +
    geom_point(alpha = 0.5) +
    #geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
    geom_smooth(method="lm", formula = y~x+I(x^2), se = FALSE )+
    scale_size_continuous(guide = "none") +
    scale_y_continuous(trans = fourth_root_power) +
    facet_wrap(~cell_type) +#, scales = "free") +
    ylab(paste("Log change in biomass over", vectors$timespan[1], "years")) +
    xlab(paste("Distance traveled to stay within critical range of", j,"")) +
    scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
    # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.2)) + 
    ggtitle(paste(survey, "mature", species))
})
mature_plot2
```

@@

mature_plot3 <- lapply(variable_names, function(j) { 
  ggplot(data, aes(x = source_density, y = log_change, 
          colour = UQ(rlang::sym(paste0("end_", j))),
    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm", se = FALSE) +
  facet_wrap(~cell_type) + #, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  scale_y_continuous(trans = sqrt) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "mature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
})
mature_plot3
```

@@

mature_plot4 <- lapply(variable_names, function(j) {  
ggplot(filter(data, distance < 10), aes(source_density, y = log_change , 
    colour = UQ(rlang::sym(paste0("end_", j))),
    alpha = 1/(distance + input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm", aes(group = cell_type)) +
  facet_wrap(~cell_type) +#, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  #scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "mature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
})
mature_plot4
```

mature_plot5 <- lapply(variable_names, function(j) {  
  ggplot(data, aes(
      x = UQ(rlang::sym(paste0(j, ".units_per_decade"))) , 
      y = (end_density - start_density), 
      colour = cell_type, 
      size = start_density)) +
    geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
    #scale_size_continuous(guide = "none") +
    #scale_y_continuous(trans = fourth_root_power) +
    ylab(paste("kg/ha change in biomass over", vectors$timespan[1], "years")) +
    xlab(paste("Change in", j," in units per decade")) +
    scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
    ggtitle(paste(survey, "mature", species))
})
mature_plot5
```

@@

```{r}
# test <- lmerTest::lmer(log(end_density) ~ cell_type * std_source_density + cell_type * distance + 
#             std_end_climate + std_end_climate2 + 
#             (1 | vector_id), 
#  #   offset = log(start_density), 
#  #   family = Gamma(link = "inverse")
#      data = data)
# hist(data$end_density)
# hist(data$log_change)
data <- vocc_by_sp[["adult"]]

model_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * std_source_density2 + cell_type * distance ") # + cell_type * distance2

@@

library(sjPlot)
#library(sjlabelled)
#library(sjmisc)
sjPlot::plot_model(test, title = "Log biomass change")
```

@@

glimpse(coefs)
rm(other_coefs)
try({other_coefs <- readRDS(paste0("data/vocc-",climate_string,"-by-spp-lmer-coefs.RDS"))})

if (exists("other_coefs")) {
coefs <- rbind(other_coefs, coefs)
}
saveRDS(coefs, file = paste0("data/vocc-",climate_string,"-by-spp-lmer-coefs.RDS"))
```

@@

Maybe try GAM?
```{r}
# library()
# 
# model_formula <- paste0("log_change ~ cell_type + s(std_source_density) + s(distance) ") # + cell_type * distance2
# 
# for (j in (variable_names) ){
#   model_formula <- paste0("", model_formula, " + s(std.", j, ") + "
#   )
# }
# 
# model_formula <- paste0("", model_formula, "+ (1 | vector_id)")
# 
# 
# test2 <- gamm4::....(as.formula(model_formula), REML = F, data = data)
#

@@

m <- gbm::gbm(
  as.formula(gbm_formula), 
  data = data, 
  distribution = "gaussian",
  n.trees = 2000, 
  interaction.depth = 3, 
  shrinkage = 0.02
  )

# m <- gbm3::gbmt(end_density ~ start_density + source_density + end_climate + 
#   distance + cell_type + X + Y, data = data, distribution = gbm_dist("Gamma"))
```

```{r}
summary(m)
```

```{r}
plot(m,1)
plot(m,2)
plot(m,3)
plot(m,6)
if (length(variable_names) > 1 ) plot(m,7)
if (length(variable_names) > 2 ) plot(m,8)

plot(m,c(1,2))
plot(m,c(1,3))
plot(m,c(1,6))

if (length(variable_names) > 1 ) plot(m,c(1,7))

plot(m,c(2,3))
plot(m,c(2,6))
if (length(variable_names) > 1 ) plot(m,c(2,7))

plot(m,c(1,2,6))
plot(m,c(2,6,3))
plot(m,c(4,5))
```

@@

tmb_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * distance + distance2") # + cell_type * std_source_density2

@@

test <- sdmTMB(data = data,
      as.formula(tmb_formula), 
      spde = spde, # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = TRUE)
test
```

```{r}

p <- predict(test)
p$residuals <- residuals(test)
#idata$resids <- p$residuals
#glimpse(p)
ggplot(p, aes(est, residuals)) +
  geom_point(alpha=0.4) +
  ylim(-8,8) +
  geom_smooth()

@@

#for (j in seq_along(variable_names)) {

immature_plot <- lapply(variable_names, function(j) { 
ggplot(data, aes(x = UQ(rlang::sym(paste0("end_", j))), y = (end_density # - start_density
  ), colour = cell_type, size = start_density, 
  group = cell_type)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method="lm", aes(x = UQ(rlang::sym(paste0("end_", j))), y = end_density
  ), formula = y ~ x + I(x^2), inherit.aes = FALSE ) + # + I(x^3)
  scale_size_continuous(guide = "none") +
  scale_y_continuous(trans = fourth_root_power) +
  xlab(paste("Final value of", j,"for cells")) +
  ylab(paste("Final biomass")) +
  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
  ggtitle(paste(survey, "immature", species))
})
immature_plot
```

@@

immature_plot2 <- lapply(variable_names, function(j) { 
  ggplot(data, aes(x = distance, y = log_change, 
      group = cell_type, size = start_density, 
      colour = UQ(rlang::sym(paste0("end_", j))))) +
    geom_point(alpha = 0.5) +
    #geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
    geom_smooth(method="lm", formula = y~x+I(x^2) )+
    scale_size_continuous(guide = "none") +
    scale_y_continuous(trans = fourth_root_power) +
    facet_wrap(~cell_type) +#, scales = "free") +
    ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
    xlab(paste("Distance traveled to stay within critical range of", j,"")) +
    scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
    # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.2)) + 
    ggtitle(paste(survey, "immature", species))
})
immature_plot2
```

@@

immature_plot3 <- lapply(variable_names, function(j) { 
  ggplot(data, aes(x = source_density, y = log_change, 
          colour = UQ(rlang::sym(paste0("end_", j))),
    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm") +
  facet_wrap(~cell_type) + #, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  scale_y_continuous(trans = sqrt) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "immature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))
})
immature_plot3
```

```{r}
data <- vocc_by_sp[["imm"]]
data <- filter(vocc_by_sp[["imm"]], distance < 50)

>>>>

Species
```{r global_options, include=FALSE}
library(dplyr)
library(ggplot2)
# library(gfplot)
# library(gfdata)
library(sdmTMB)
library(gfranges)
# 
# knitr::opts_chunk$set(
#   fig.width = 11, fig.height = 8.5,
#   echo = FALSE, warning = FALSE, message = FALSE
# )
options(scipen = 999)
options(digits = 2) 
```

```{r}
# # Species run so far...
# species <- "Arrowtooth Flounder"
# species <- "Pacific Cod"
# species <- "Sablefish"
 species <- "Silvergray Rockfish"
# species <- "Lingcod"
# species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
# species <- "Quillback Rockfish"
# species <- "Pacific Ocean Perch"
# species <- "Yelloweye Rockfish"
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
```

Region 
```{r}
# region <- "All synoptic surveys"
# region <- "West Coast Vancouver Island"
 region <- "Both odd year surveys"
# region <- "West Coast Haida Gwaii"
```

@@

#climate <- "temperature"
temp_threshold <- c(0.5)

#climate <- "do"
do_threshold <- c(0.1)

climate <- "do and temperature"
max_thresholds <- c(Inf, temp_threshold)
min_thresholds <- c(do_threshold, Inf)
```

Get most up to date optimal values from mountain plots for most abundant year

```{r}
params <- readRDS(paste0("data/optimal-values-by-species.rds"))
params <- params[rev(order(as.Date(params$date))), ]
params <- params[!duplicated(params$species),]

optimal_do <- round(params[params$species == species, ]$optimal_do, digits = 2)
optimal_temp <- round(params[params$species == species, ]$optimal_temp, digits = 2)

optimal_values <- c(optimal_do, optimal_temp)
```

@@

Run all subsequent code...
```{r, message=FALSE, warning=FALSE}
if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  trend_indices_do <- c(1, 1, 1, 2, 2)
  trend_indices_temp <- c(1, 1, 1, 2, 2)
  years <- NULL
}

@@

```{r}
rm(vocc)
try({
  vocc <- readRDS(paste0("data/vocc-", ssid_string,"-", climate_string, "-",
    min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
})

@@

saveRDS(vocc, file = paste0("data/vocc-", ssid_string,"-", climate_string, "-",
    min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
}
# glimpse(vocc)
```

### Plot all climate vectors
```{r eval = FALSE}

@@

Filter vectors to match optimal range for chosen species
```{r}
vectors <- trim_vector_data(vocc, variable_names, optimal_values, min_thresholds, max_thresholds)

if (length(vectors[ ,2]) == 0) {
  other_coefs <- readRDS(paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
  coefs <- other_coefs[1,]
  coefs$var <- "NA"
  coefs$beta <- "NA"
  coefs$se <- "NA"
  coefs$model <- "NA"
  coefs$species <- spp 
  coefs$lifestage <- "NA"
  coefs$climate <- climate 
  coefs$min_thresholds <- min_string
  coefs$max_thresholds <- max_string
  coefs$optimum <- optimal_string
  coefs$start_time <- start_time
  coefs$timespan <- delta_t_total
  coefs$region <- region
  coefs$ssid_string <- ssid_string

  coefs <- rbind(other_coefs, coefs) %>% distinct()
  saveRDS(coefs, file = paste0("data/", climate_string,"-vocc-by-spp-sdmTMB-coefs.RDS"))

### Plot remaining climate vectors for regions where change exceeds thresholds
```{r}

if(climate== "temperature") {
vocc_plot <- plot_vocc(vectors,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  white_zero = TRUE,
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
)
vocc_plot + ggtitle(climate)
}

if (climate == "do") {
vocc_plot <- plot_vocc(vectors,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
)
vocc_plot + ggtitle(climate)
}

vocc_plot_temp <- plot_vocc(vectors,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "temp_diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  white_zero = TRUE,
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
)+ ggtitle("temperature")

vocc_plot_do <- plot_vocc(vectors,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "do_diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
) + ggtitle("do")

@@

# Extract biomass values to match spatial and temporal range of climate vectors

@@

biomass_change <- make_trend_data(i,
  ssid = model_ssid,
  start_time = start_time,
  skip_time = skip_time,
  end_time = end_time,
  input_cell_size = input_cell_size,
  scale_fac = scale_fac,
  delta_t_total = delta_t_total,
  delta_t_step = delta_t_step,
  indices = indices,
  variable_names = "est_exp"
)
#browser()
vectors <- vectors %>%
      mutate(
        start_cell = icell,
        vector_id = paste0(id, "_to_", tid),
        distance = ifelse(distance>120, 120, distance),
        distance2 = distance*distance,
        source_end_value = e, # may need to turn into numeric vector later...
        target_value = target_values
      )

@@

# combine source and target cell data
vocc_by_sp <- rbind(vector_source, vector_target) %>% 
  mutate(cell_type = as.factor(cell_type)) %>%
  filter(!is.na(icell)) %>% 
  filter(!is.na(source_density)) %>% filter(!is.na(end_density)) %>%

  # remove potential source cells with lowest 10% quantile of fish biomass  
  filter(source_density > quantile(source_density, 0.1)) %>%

  # calculate percent change in biomass
  mutate(
    log_change = log(end_density/start_density),
    perc_change = end_density/start_density
    )
#browser()
end_climate <- purrr::map_df(strsplit(vocc_by_sp$end_climate, " "), 
                   function(x) data.frame(t(as.numeric(x))))

@@

vocc_by_sp
})
```

@@

Check final climate-biomass relationship
```{r}
data <- vocc_by_sp[["adult"]]

mature_plot <- lapply(variable_names, function(j) { 
ggplot(data, aes(x = UQ(rlang::sym(paste0("end_", j))), y = end_density, colour = cell_type, size = start_density, 
  group = cell_type)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method="lm", aes(x = UQ(rlang::sym(paste0("end_", j))), y = end_density),
    formula = y ~ x + I(x^2), inherit.aes = FALSE) + # + I(x^3)
  # scale_size_continuous(guide = "none") +
  scale_y_continuous(trans = log10) +
  xlab(paste("Final value of", j,"for cells")) +
  ylab(paste("Final biomass")) +
  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.7)) + 
  ggtitle(paste(survey, "mature", species))
})
print(mature_plot)
```

Check distance threshold of 50 looks reasonable
```{r}
data <- vocc_by_sp[["adult"]]
#data <- filter(vocc_by_sp[["adult"]],  distance < 90) # distance > 1 &

mature_plot2 <- lapply(variable_names, function(j) { 
  ggplot(data, aes(x = distance, y = log_change, 
      group = cell_type, size = start_density, 
      colour = UQ(rlang::sym(paste0("end_", j))))) +
    geom_point(alpha = 0.5) +
    #geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
    geom_smooth(method="lm",  se = FALSE )+ # formula = y~x+I(x^2),
    scale_size_continuous(guide = "none") +
    facet_wrap(~cell_type) +#, scales = "free") +
    ylab(paste("Log change in biomass over", vectors$timespan[1], "years")) +
    xlab(paste("Distance traveled to stay within critical range of", j,"")) +
    scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
    # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.2)) + 
    ggtitle(paste(survey, "mature", species))
})
mature_plot2
```

@@

mature_plot3 <- lapply(variable_names, function(j) { 
  ggplot(data, aes(x = source_density, y = log_change, 
          colour = UQ(rlang::sym(paste0("end_", j))),
    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm", se = FALSE) +
  facet_wrap(~cell_type) + #, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  #scale_y_continuous(trans = sqrt) +
  ylab(paste("Log change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "mature", species))
})
mature_plot3
```

@@

mature_plot4 <- lapply(variable_names, function(j) {  
ggplot(filter(data, distance < 10), aes(source_density, y = log_change , 
    colour = UQ(rlang::sym(paste0("end_", j))),
    alpha = 1/(distance + input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm", aes(group = cell_type)) +
  facet_wrap(~cell_type) +#, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  #scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("Log change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "mature", species))
})
mature_plot4
```

@@

```{r}
data <- vocc_by_sp[["adult"]] #%>% filter(distance < 50)

model_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * distance") # + cell_type * std_source_density2 + cell_type * distance2") #

@@

#library(sjPlot)
#library(sjlabelled)
#library(sjmisc)
sjPlot::plot_model(test, title = paste("Log adult", species, "biomass change"))
```

@@

rm(other_coefs)
try({other_coefs <- readRDS(paste0("data/vocc-",climate_string,"-by-spp-lmer-coefs.RDS"))})

if (exists("other_coefs")) {
coefs <- rbind(other_coefs, coefs) %>% distinct() %>% 
 mutate_if(is.numeric, funs(signif(.,digits = 2)))
}

coefs <- coefs %>% distinct()

saveRDS(coefs, file = paste0("data/vocc-",climate_string,"-by-spp-lmer-coefs.RDS"))
```

@@

Maybe try GAM?
```{r}
# library()
# 
# model_formula <- paste0("log_change ~ cell_type + s(std_source_density) + s(distance) ") # + cell_type * distance2
# 
# for (j in (variable_names) ){
#   model_formula <- paste0("", model_formula, " + s(std.", j, ") + "
#   )
# }
# 
# model_formula <- paste0("", model_formula, "+ (1 | vector_id)")
# 
# 
# test2 <- gamm4::....(as.formula(model_formula), REML = F, data = data)
# 
```

@@

mboost <- gbm::gbm(
  as.formula(gbm_formula), 
  data = data, 
  distribution = "gaussian",
  n.trees = 2000, 
  interaction.depth = 3, 
  shrinkage = 0.02
  )
```

```{r}
summary(mboost)
```

```{r}
plot(mboost,3)

plot(mboost,c(1,6,3))
if (length(variable_names) > 1 ) plot(mboost,c(1,7,3))
if (length(variable_names) > 2 ) plot(mboost,c(1,8,3))

plot(mboost,c(2,6,3))
if (length(variable_names) > 1 ) plot(mboost,c(2,7,3))
if (length(variable_names) > 2 ) plot(mboost,c(2,8,3))
```

Other gbm plots
```{r eval=FALSE}
plot(mboost, 1)
plot(mboost,2)
plot(mboost,6)
if (length(variable_names) > 1 ) plot(mboost,7)
if (length(variable_names) > 2 ) plot(mboost,8)

plot(mboost,c(1,2))
plot(mboost,c(1,3))
plot(mboost,c(1,6))

if (length(variable_names) > 1 ) plot(mboost,c(1,7))
if (length(variable_names) > 2 ) plot(mboost,c(1,8))

plot(mboost,c(2,3))
plot(mboost,c(2,6))
if (length(variable_names) > 1 ) plot(mboost,c(2,7))
if (length(variable_names) > 2 ) plot(mboost,c(2,8))

plot(mboost,c(4,5))
```

@@

tmb_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * distance") # + cell_type * distance2 cell_type * std_source_density2 +

@@

mtmb <- sdmTMB(data = data,
      as.formula(tmb_formula), 
      spde = spde, # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = TRUE)
mtmb 
```

```{r}
p <- predict(mtmb)
p$residuals <- residuals(mtmb)
#idata$resids <- p$residuals
#glimpse(p)
ggplot(p, aes(est, residuals)) +
  geom_point(alpha=0.4) +
  ylim(-8,8) +
  geom_smooth()

@@

immature_plot <- lapply(variable_names, function(j) { 
ggplot(data, aes(x = UQ(rlang::sym(paste0("end_", j))), y = (end_density), 
    colour = cell_type, size = start_density, 
    group = cell_type)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method="lm", aes(x = UQ(rlang::sym(paste0("end_", j))), y = end_density),
    formula = y ~ x + I(x^2), inherit.aes = FALSE ) + # + I(x^3)
  scale_size_continuous(guide = "none") +
  scale_y_continuous(trans = log10) +
  xlab(paste("Final value of", j,"for cells")) +
  ylab(paste("Final biomass")) +
  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
  ggtitle(paste(survey, "immature", species))
})
immature_plot
```

@@

immature_plot2 <- lapply(variable_names, function(j) { 
  ggplot(data, aes(x = distance, y = log_change, 
      group = cell_type, size = start_density, 
      colour = UQ(rlang::sym(paste0("end_", j))))) +
    geom_point(alpha = 0.5) +
    #geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
    geom_smooth(method="lm") + #, formula = y~x+I(x^2) 
    scale_size_continuous(guide = "none") +
    facet_wrap(~cell_type) +#, scales = "free") +
    ylab(paste("Log change in biomass over", vectors$timespan[1], "years")) +
    xlab(paste("Distance traveled to stay within critical range of", j,"")) +
    scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
    # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.7)) + 
    ggtitle(paste(survey, "immature", species))
})
immature_plot2
```

@@

immature_plot3 <- lapply(variable_names, function(j) { 
  ggplot(data, aes(x = source_density, y = log_change, 
          colour = UQ(rlang::sym(paste0("end_", j))),
    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm") +
  facet_wrap(~cell_type) + #, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  #scale_y_continuous(trans = sqrt) +
  ylab(paste("Log change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "immature", species))
})
immature_plot3
```
